CF Registry Analysis

### 2024-07-30 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
vertex terms added

```{r}
#| message: false
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_perc_cat, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical")
        ) %>%
        add_p() %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
screening <- annual %>%
  group_by(ReviewYear, age_group) %>%
  summarise(
    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose)),
    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
    .groups = "drop"
  ) %>%
  pivot_longer(`Screened HbA1c`:`Screened Either`)
screening_plot <- ggplot(screening, aes(x = ReviewYear, y = value, color = age_group)) +
  geom_line() +
  xlab("Year") +
  facet_wrap(~name) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_discrete(name = "Age Group") +
  theme(text = element_text(size = 20), axis.title.y = element_blank())
ggsave("./Dissemination/Figures/screening_rates.png",
  plot = screening_plot, width = 12, height = 9, units = "in"
)
```

@@

```{r results='asis'}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Make table for each variable
age <- ogtts %>%
  select(`Age Group`, OGTTs) %>%
  tbl_summary(by = `Age Group`) %>%
  add_p() %>%
  print()
gender <- ogtts %>%
  select(Gender, OGTTs) %>%
  tbl_summary(by = Gender) %>%
  add_p() %>%
  print()
muts <- ogtts %>%
  select(MutClass, OGTTs) %>%
  tbl_summary(by = MutClass) %>%
  add_p() %>%
  print()
race <- ogtts %>%
  select(Race, OGTTs) %>%
  tbl_summary(by = Race) %>%
  add_p() %>%
  print()
eth <- ogtts %>%
  select(Hispanicrace, OGTTs) %>%
  tbl_summary(by = Hispanicrace) %>%
  add_p() %>%
  print()
```

@@

## Cox PH model

@@

@@

```{r results='asis'}
invisible(lapply(xvars, function(v) {
  cat("\n")
  cat("#### By group")
  cat("\n")
  annual %>%
    group_by_at(v) %>%
    summarise(
      n = length(unique(eDWID)),
      `CFRD+` = sum(CFRD == "CFRD+"),
      .groups = "drop"
    ) %>%
    gt() %>%
    print()
  cat("\n")
  cat("#### By and year")
  cat("\n")
  annual %>%
    group_by_at(c(v, "ReviewYear")) %>%
    summarise(
      n = length(unique(eDWID)),
      `CFRD+` = sum(CFRD == "CFRD+"),
      .groups = "drop"
    ) %>%
    gt() %>%
    print()
  cat("\n")
}))
```

>>>>

```{r}
#| message: false
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_perc_cat, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical")
        ) %>%
        add_p() %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
screening <- annual %>%
  group_by(ReviewYear, age_group) %>%
  summarise(
    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose)),
    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
    .groups = "drop"
  ) %>%
  pivot_longer(`Screened HbA1c`:`Screened Either`)
screening_plot <- ggplot(screening, aes(
  x = ReviewYear, y = value,
  color = age_group
)) +
  geom_line() +
  xlab("Year") +
  facet_wrap(~name, nrow = 3) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_discrete(name = "Age Group") +
  theme(text = element_text(size = 20), axis.title.y = element_blank())
screening_plot
ggsave("./Dissemination/Figures/screening_rates.png",
  plot = screening_plot, width = 12, height = 9, units = "in"
)
```

@@

```{r results='asis'}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Make table for each variable
age <- ogtts %>%
  select(`Age Group`, OGTTs) %>%
  tbl_summary(by = `Age Group`) %>%
  add_ci() %>%
  add_p() %>%
  print()
gender <- ogtts %>%
  select(Gender, OGTTs) %>%
  tbl_summary(by = Gender) %>%
  add_ci() %>%
  add_p() %>%
  print()
muts <- ogtts %>%
  select(MutClass, OGTTs) %>%
  tbl_summary(by = MutClass) %>%
  add_ci() %>%
  add_p() %>%
  print()
race <- ogtts %>%
  select(Race, OGTTs) %>%
  tbl_summary(by = Race) %>%
  add_ci() %>%
  add_p() %>%
  print()
eth <- ogtts %>%
  select(Hispanicrace, OGTTs) %>%
  tbl_summary(by = Hispanicrace) %>%
  add_ci() %>%
  add_p() %>%
  print()
```

@@

## Cox PH models

@@

### Vx770

Years prior to 2011 were excluded from the following model

```{r results='asis'}
annual %>%
  group_by(ReviewYear) %>%
  count(A_Vx770) %>%
  gt()
```

```{r results='asis'}
f_vx770 <- update(f, . ~ . + A_Vx770)
vx770_cox_res <- coxph(f_vx770,
  data = annual[annual$ReviewYear >= 2011, ],
  cluster = eDWID
)
vx770_cox_res %>%
  tbl_regression()
ggcoef_model(vx770_cox_res,
  show_p_values = F, signif_stars = F, colour = F
)
```

### Vx809comb

Years prior to 2015 were excluded from the following model due to missing data.

```{r results='asis'}
annual %>%
  group_by(ReviewYear) %>%
  count(A_VX809comb) %>%
  gt()
```

```{r}
f_vx809comb <- update(f, . ~ . + A_VX809comb)
vx809comb_cox_res <- coxph(f_vx809comb,
  data = annual[annual$ReviewYear >= 2015, ],
  cluster = eDWID
)
vx809comb_cox_res %>%
  tbl_regression()
ggcoef_model(vx809comb_cox_res,
  show_p_values = F, signif_stars = F, colour = F
)
```

@@

```{r results='asis'}
invisible(lapply(c(xvars, "A_Vx770", "A_VX809comb"), function(v) {
  cat("\n")
  cat("#### By group")
  cat("\n")
  annual %>%
    group_by_at(v) %>%
    summarise(
      n = length(unique(eDWID)),
      `CFRD+` = sum(CFRD == "CFRD+"),
      .groups = "drop"
    ) %>%
    gt() %>%
    print()
  cat("\n")
  cat("#### By and year")
  cat("\n")
  annual %>%
    group_by_at(c(v, "ReviewYear")) %>%
    summarise(
      n = length(unique(eDWID)),
      `CFRD+` = sum(CFRD == "CFRD+"),
      .groups = "drop"
    ) %>%
    gt() %>%
    print()
  cat("\n")
}))
```
